Switch equivalence and weak prime domains for fusions

### 2024-06-03 fusione.tex
minori

\begin{definition} 
	Given a set $I$, a  \emph{source}  (\emph{sink}) is a family $\{f_i\}_{i\in I}$ of arrows  $f_i\colon X\to Y_i$ ($f_i\colon Y_i\to X$) with the same (co)domain. A $\emph{wide pushout}$ (pullback) is the colimit (limit) of a source (sink). We will use $c_i$ ($p_i$) to denote the coprojection from $Y_i$ (the projection to $Y_i$) and $c_X$ to denote the one from $X$.
\end{definition}

\begin{remark}\label{rem:copro}
	Given a wide pushout $(C, \{c_i\}_{i\in I\cup \{X\}})$ on a source $\{f_i\}_{i\in I}$ with $f_i\colon X\to Y_i$, the coprojection $c_X$ is such that, for every $i\in I$, the following diagram commute
	\[\xymatrix{&X\ar[dl]_{f_i} \ar[dr]^{c_X}\\
		Y_i\ar[rr]_{c_{i}} && C}\]
\end{remark}

\begin{proposition} \label{prop:mp} For every proper factorization system $(\mathcal{E}, \mathcal{M})$ on a category $\X$ the following hold:
	\begin{enumerate}
		\item  for every pushout square as the one below,  $e\in \mathcal{E}$ implies $n\in \mathcal{E}$ 
		\[\xymatrix{X\ar[r]^{g} \ar[d]_{e} & V \ar[d]^{n} \\ Y \ar[r]_{f}  & Z}\] 
		\item  if $(C, \{c_i\}_{i\in I\cup \{X\}})$ is a wide pushout on a source $\{e_i\}_{i\in I}$ such that $e_i\colon X\to Y_i$ is in $\mathcal{E}$ for every $i\in I$, then every coprojection is in $\mathcal{E}$ too. 
	\end{enumerate}
\end{proposition}
\begin{proof}
	\begin{enumerate}
		\item

		\item By \cref{rem:copro} and point three of \cref{cor:iso} it is enough to show that $c_X$ is in $\mathcal{E}$. Since $(\mathcal{E}, \mathcal{M})$ is a factorization system then $c_X=m\circ e$ for some $m\colon E\to C$ in $\mathcal{M}$ and $e\colon X\to E$ in $\mathcal{E}$ to get, from \cref{rem:copro}, a square
		\[\xymatrix{X\ar[r]^{e} \ar[d]_{e_i} & V \ar[d]^{m} \\ Y_i \ar@{.>}^{k_i}[ur] \ar[r]_{c_i}  & Z}\]
		and the left lifting property provides, for every $i\in I$, the dotted arrow $k_i\colon Y_i\to e$. Let $k$ be the the induced arrow $Z\to V$. Then
		\begin{align*}
			m\circ k\circ c_i&=m\circ k_i\\&=c_i
		\end{align*}
		hence $m\circ k=\id{Z}$. By \cref{cor:iso} $m\in \mathcal{M}$ thus it is an isomorphism and we can conclude. \qedhere 
	\end{enumerate}
\end{proof}

	When $\mathcal{M}$ coincides with the class of all monos we will suppress the prefix, so that a \emph{subobject} is just an equivalence class of a mono and $((\msub{X}{X}), \leq)$ will denote the ordered class of such subobjects of $X$.
\end{definition}
\begin{remark}\label{rem:size}
	Every $\mathcal{M}$-adhesive cateory $\X$ admits a full and faithful functor $F:\X\to \E$ into a topos $\E$, sending arrows in $\mathcal{M}$ to monomorphisms \cite{castelnovo2023thesis}. Since every topos is \emph{well-powered}, i.e. $\msub{E}{E}$ is a set for every object $E$ \cite{maclane2012sheaves}, we can deduce  that, for every $X$ in $\X$, $\sub{M}{X}{X}$ is actually a set too.
\end{remark}

\begin{remark}
	contenuto...
\end{remark}

Our next step is to compute infima in $(\sub{M}{X}{X}, \leq)$.

\begin{proposition}
	inf
\end{proposition}
\begin{proof}
\end{proof}
\begin{corollary}
	contenuto...
\end{corollary}

@@

\subsection{Left-linear DPO-rewriting systems}
We are now going to study rewriting systems in $\mathcal{M}, \mathcal{N}$-adhesive categories.

@@

>>>>

@@

\subsection{Left-linear DPO-rewriting systems}
We are now going to study rewriting systems in $\mathcal{M}$-adhesive categories.

@@

\subsubsection{$\mathcal{M}$-subobjects and factorization systems}

	When $\mathcal{M}$ coincides with the class of all monos we will suppress the prefix, so that a \emph{subobject} is just an equivalence class of a mono and $(\msub{X}{X}, \leq)$ will denote the ordered class of such subobjects of $X$.
\end{definition}

\begin{remark}\label{rem:pbo} Let $f\colon X\to Y$ be an arrows admitting a pullback along arrows in $\mathcal{M}/Y$. The categories associated to $(\sub{M}{X}{X}, \leq)$ and $(\sub{M}{Y}{Y}, \leq)$ are equivalent to $\mathcal{M}/X$ and $\mathcal{M}/Y$, respectively, thus we get a monotone function $f^{-1}\colon \sub{M}{X}{Y}\to \sub{M}{X}{X}$ sending $[m]$ to $[f^*(m)]$.
\end{remark}

\begin{remark}\label{rem:size}
	Every $\mathcal{M}$-adhesive cateory $\X$ admits a full and faithful functor $F:\X\to \E$ into a topos $\E$, sending arrows in $\mathcal{M}$ to monomorphisms \cite{castelnovo2023thesis}. Since every topos is \emph{well-powered}, i.e. $\msub{E}{E}$ is a set for every object $E$ \cite{maclane2012sheaves}, we can deduce  that, for every $X$ in $\X$, $\sub{M}{X}{X}$ is actually a set too.
\end{remark}

Our next step is to compute finite infima in $(\sub{M}{X}{X}, \leq)$.

\begin{proposition}\label{prop:inf}
	Let $\mathcal{M}$ be a class of monos  in a category $\X$.  Let  $m\colon M\to X$ be an arrow in $\mathcal{M}$, then for every family $\{m_i\}_{i\in I}$ of sobjects in $\mathcal{M}/X$ the following are equivalent:
	\begin{enumerate}
		\item $[m]$ is the infimum of $\{[m_i]\}_{i\in I}$;
		\item there is a limiting cone $(m, \{k_i\}_{i\in I})$ for the diagram in $\mathcal{M}/X$ defined by the family $\{m_i\}_{i\in I}$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	$(1) \Rightarrow (2)$ Let us denote by $M_i$ the domain of $m_i$. By hypothesis $m\leq m_i$, so that there exists $k_i\colon M\to M_i$. Let $(n, \{h_i\}_{i\in I})$ be another cone in $\mathcal{M}/X$, that is we have arrows $n\colon N\to X$ in $\mathcal{M}$ and $h_i\colon N\to M_i$. In particular $[n]$ is a lower bound in $(\sub{M}{X}{X}, \leq)$ for $\{[m_i]\}_{i\in I}$, thus there exists a, necessarily unique, arrow $t\colon N\to M$ such that $m\circ t=n$. Now, for every $i\in I$ we have
	\begin{align*}
		m_i\circ k_i\circ t&=m\circ t \\&= n\\&=m_i\circ h_i
	\end{align*}
	Since $m_i$ is mono we conclude that $k_i\circ t=h_i$.

	\smallskip \noindent
	$(2) \Rightarrow (1)$ The family $\{k_i\}_{i\in I}$ of coprojections witnesses that $[m]$ is a lower bound in $(\sub{M}{X}{X}, \leq)$ for $\{[m_i]\}_{i\in I}$. Let $[n]$ be another one, thus we have a cone $(n, \{h_i\}_{i\in I})$ in $\mathcal{M}/X$ and so a unique $t:N\to M$, where $N$ is the domain of $n$, such that $k_i\circ t=h_i$. Thus
	\begin{align*}
		m\circ t &= m_i\circ k_i\circ t\\&=m_i\circ h_i\\&=n
	\end{align*}
	We can thus conclude that $[n]\leq [m]$.
\end{proof}
\begin{corollary}
Let $\mathcal{M}$ be a class of monos in a category $\X$, then the following are true:
\begin{enumerate}
	\item if $\mathcal{M}$ contains $\id{X}$ then $(\sub{M}{X}{X}, \leq)$ has a maximum;
	\item if $\mathcal{M}$ is stable under pullback, closed under composition and $\X$ has pullbacks of arrows in $\mathcal{M}$ along arrows in $\mathcal{M}$ then $(\sub{M}{X}{X}, \leq)$ has binary infima.
\end{enumerate}
\end{corollary}

We end this appendix showing how to construct a factorization system given a suitable class of monos $\mathcal{M}$.

\todo{fact system}
